Mining Interaction Log Data in a Creative Arts Practice using Transition Matrices

### 2015-09-24 flux-entropy-paper/flux-entropy-paper.tex
now fits on 4 pages phew!

While the interaction sessions recorded in our corpus were performed
by many different participants with varying experience levels, in
different contexts and on different instrumental setups, the common
interface and method of recording the data allows all the sessions to
be transformed into gesture sequences, and transition matrices. While
these matrices can be used to visualise the structure of individual
performances, as has been done in ~\cite{Swift:2014tya}, we also wish to extract quantitative
measures from these TMs to compare the large number of sessions in our
corpus.
To do this we calculate two high-level quantities from each session:
``flux'' and ``entropy''. The flux of a sequence is defined as the
ratio of state transitions (e.g. $A \rightarrow B$ where $A \neq B$),
to self transitions (e.g. $A \rightarrow A $). The flux of a TM $P$ is
given by: $\mathrm{flux}(P) = 1 - \mathrm{tr}(P)$ where the trace
$\mathrm{tr}(P)$ is the sum of the diagonal entries of $P$.
Intuitively, flux is a measure of how frequently the participant/group
changes state. Flux returns a value in the interval $[0,1]$ and will
return 0 when participants never change state, and 1 when participants
never stay in the same state for two measurements in a row.

@@

%\pagebreak
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{references}

>>>>

While the interaction sessions recorded in our corpus were performed
by many different participants in
different contexts and on different instrumental setups, the common
interface and method of recording the data allows all the sessions to
be transformed into gesture sequences, and transition matrices. While
these matrices can be used to visualise the structure of individual
performances, as has been done in ~\cite{Swift:2014tya}, we also wish to extract quantitative
measures from these TMs to compare the large number of sessions in our
corpus.
To do this we calculate two high-level quantities from each session:
``flux'' and ``entropy''. The flux of a sequence is defined as the
ratio of state transitions (e.g. $A \rightarrow B$ where $A \neq B$),
to self transitions (e.g. $A \rightarrow A $). The flux of a TM $P$ is
given by: $\mathrm{flux}(P) = 1 - \mathrm{tr}(P)$ where the trace
$\mathrm{tr}(P)$ is the sum of the diagonal entries of $P$.
Intuitively, flux is a measure of how frequently the participant/group
changes state. Flux returns a value in the interval $[0,1]$ and will
return 0 when participants never change state, and 1 when participants
never stay in the same state for two measurements in a row.

@@

\pagebreak
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{references}
